Limitations on the Two-Infall Scenario in the Era of Large Stellar Age Catalogs

### 2025-04-03 src/tex/ms.tex
Updates from Overleaf

The population-averaged nucleosynthetic yields of CCSNe are still fairly uncertain. This problem is exacerbated by the complexity of the CCSN explosion landscape \citep{sukhbold_core-collapse_2016}. Recently, \citet{weinberg_scale_2024} used a measurement of the mean Fe yield of CC SNe by \citet{rodriguez_iron_2023} and the plateau in stellar \aFe abundances at low metallicity to infer population-averaged yields of $y/Z_\odot\approx1$\textemdash in other words, the scale of SN yields is similar to the Solar abundance. However, \citet{johnson_milky_2024} found that GCE models with yields at this scale approach present-day abundances too slowly to match the observed age--metallicity relation. Previous multi-zone models using {\tt VICE} \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024} adopted higher yields based on \citet{chieffi_explosive_2004} and \citet{limongi_nucleosynthesis_2006} ($y/Z_\odot\approx2.6$).

We investigate yield sets at multiple scales of the Solar abundance. The CCSN yield of O is directly set by the Solar scale: $y_{\rm O}^{\rm CC}=(y/Z_\odot)Z_{\rm O,\odot}$. For Fe, the CCSN yield is set by the \aFe ``plateau'' at low metallicity, $\mathaFe_{\rm CC}$, such that $y_{\rm Fe}^{\rm CC}=(y/Z_\odot) Z_{\rm Fe,\odot} 10^{-\mathaFe}$ \citep[for further discussion on the empirical yield scale and the CCSN plateau, see][]{weinberg_scale_2024}. We adopt the \citet{asplund_chemical_2009} Solar abundances: $Z_{\rm O,\odot}=5.72\times10^{-3}$ and $Z_{\rm Fe,\odot}=1.29\times10^{-3}$. Our yield sets are presented in Table \ref{tab:yields}.

The SN Ia yield of Fe, $y_{\rm Fe}^{\rm Ia}$, is calculated so that the combined Fe yield $y_{\rm Fe}^{\rm Ia+CC}=y/Z_\odot$. We further increase $y_{\rm Fe}^{\rm Ia}$ by a factor of 30\% and 10\% for the \yZ{1} and \yZ{2} yield sets, respectively, to ensure that our models reach $\mathOFe\approx0.0$ by $t=13.2\,{\rm Gyr}$. The fifth row of Table \ref{tab:yields} reports the integrated SN Ia rate $N_{\rm Ia}/M_\star$ from each yield set, assuming a mean Fe yield per SN Ia of $\overline m_{\rm Fe}^{\rm Ia}=0.7$ M$_\odot$ \citep{mazzali_common_2007,howell_effect_2009}. The rate for the \yZ{1} yield set is slightly higher than the volumetric rate of $N_{\rm Ia}/M_\star=(1.3\pm0.1)\times10^{-3}\,{\rm M}_\odot^{-1}$ reported by \citet{maoz_star_2017}, but is consistent with their measurement of $N_{\rm Ia}/M_\star=(1.6\pm0.3)\times10^{-3}\,{\rm M}_\odot^{-1}$ for field galaxies. The rate for the \yZ{2} yield set is consistent with the measurement of $N_{\rm Ia}/M_\star=(2.2\pm1.0)\times10^{-3}\,{\rm M}_\odot^{-1}$ by \citet{maoz_type-ia_2012}, while the rate for the \yZ{3} yield set is generally higher than literature values.

@@

\todo{I'd like to compare these yields to previous studies, but it's been very difficult to make that comparison. I'm unsure whether the \citet{francois_evolution_2004} yields are net or gross (probably gross since their baseline is \citet{woosley_evolution_1995}). I tried to implement their yield tables in VICE, but the chemical evolution tracks came out looking very wrong. Also, they use a weird IMF.} \citet{francois_evolution_2004} report that their models are insensitive to changes in the yield at a factor of 2 level.

\begin{figure}
    \centering
    \includegraphics{figures/yield_outflow.pdf}
    \caption{The abundance evolution of three one-zone models with different yield and outflow settings. Table \ref{tab:yields} presents the population-averaged yields for each model. The gray points with error bars indicate the mode of the APOGEE abundance distributions from the Solar neighborhood ($7\leq R_{\rm gal}<9\,{\rm kpc}$, $0\leq|z|<0.5\,{\rm kpc}$) in 1 Gyr-wide bins of the \citet{leung_variational_2023} age estimates.}
    \label{fig:yield-outflow}
\end{figure}

Figure \ref{fig:yield-outflow} illustrates the effect of the yield scaling on the abundance evolution in one-zone models. All models feature a rapid dilution of the ISM metallicity by $\sim0.5-0.8$ dex, visible in the top two panels, brought on by the infall of pristine gas at $t_{\rm max}$. For the model with $y/Z_\odot=1$, this dilution persists for some time and the metallicity does not return to Solar until the present day. The models with higher yields and outflows recover from this dilution more quickly, returning to Solar metallicity by $\sim5\,{\rm Gyr}$ ago. However, the high-yield models experience a decline in [O/Fe] of $\sim0.2$ dex between the second infall and the present day, contrasted with the smaller decline of $\sim0.1$ dex in the low-yield model.

>>>>

The population-averaged nucleosynthetic yields of CCSNe are still fairly uncertain. This problem is exacerbated by the complexity of the CCSN explosion landscape \citep{sukhbold_core-collapse_2016}. Recently, \citet{weinberg_scale_2024} used a measurement of the mean Fe yield of CC SNe by \citet{rodriguez_iron_2023} and the plateau in stellar \aFe abundances at low metallicity to infer population-averaged yields of $y/Z_\odot\approx1$\textemdash in other words, the scale of SN yields is similar to the Solar abundance. However, \citet{johnson_milky_2024} found that GCE models with yields at this scale approach present-day abundances too slowly to match the observed age--metallicity relation. Previous multi-zone models using {\tt VICE} \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024} adopted higher yields ($y/Z_\odot\approx2.6$) based on \citet{chieffi_explosive_2004} and \citet{limongi_nucleosynthesis_2006}; however, in order to produce a realistic evolution of [O/Fe], those studies adopted an integrated SN Ia rate which is high compared to the measurement of \citet{maoz_star_2017}.

We investigate yield sets at multiple scales of the Solar abundance. The CCSN yield of O is directly set by the Solar scale: $y_{\rm O}^{\rm CC}=(y/Z_\odot)Z_{\rm O,\odot}$. For Fe, the CCSN yield is set by the \aFe ``plateau'' at low metallicity, $\mathaFe_{\rm CC}$, such that $y_{\rm Fe}^{\rm CC}=(y/Z_\odot) Z_{\rm Fe,\odot} 10^{-\mathaFe}$ \citep[for further discussion on the empirical yield scale and the CCSN plateau, see][]{weinberg_scale_2024}. We adopt the \citet{asplund_chemical_2009} Solar abundances: $Z_{\rm O,\odot}=5.72\times10^{-3}$ and $Z_{\rm Fe,\odot}=1.29\times10^{-3}$. Our yield sets are presented in Table \ref{tab:yields}. We consider \yZ{1} representative of the empirical yield scale, whereas \yZ{2-3} span a range of theoretical predictions.

The SN Ia yield of Fe, $y_{\rm Fe}^{\rm Ia}$, is calculated so that the combined Fe yield $y_{\rm Fe}^{\rm Ia}+y_{\rm Fe}^{\rm CC}=y/Z_\odot$. We further increase $y_{\rm Fe}^{\rm Ia}$ by a factor of 30\% and 10\% for the \yZ{1} and \yZ{2} yield sets, respectively, to ensure that our models reach $\mathOFe\approx0.0$ by $t=13.2\,{\rm Gyr}$. The fifth row of Table \ref{tab:yields} reports the integrated SN Ia rate 
\begin{equation}
    \frac{N_{\rm Ia}}{M_\star} = \frac{y_{\rm Fe}^{\rm Ia}}{\bar m_{\rm Fe}^{\rm Ia}}
    \label{eq:snia-rate}
\end{equation}
from each yield set, assuming a mean Fe yield per SN Ia of $\overline m_{\rm Fe}^{\rm Ia}=0.7$ M$_\odot$ \citep{mazzali_common_2007,howell_effect_2009}. The rate for the \yZ{1} yield set is slightly higher than the volumetric rate of $N_{\rm Ia}/M_\star=(1.3\pm0.1)\times10^{-3}\,{\rm M}_\odot^{-1}$ reported by \citet{maoz_star_2017}, but is consistent with their measurement of $N_{\rm Ia}/M_\star=(1.6\pm0.3)\times10^{-3}\,{\rm M}_\odot^{-1}$ for field galaxies. The rate for the \yZ{2} yield set is consistent with the measurement of $N_{\rm Ia}/M_\star=(2.2\pm1.0)\times10^{-3}\,{\rm M}_\odot^{-1}$ by \citet{maoz_type-ia_2012}, while the rate for the \yZ{3} yield set is generally higher than literature values.

@@

Many previous two-infall studies have adopted the yields of \citet{francois_evolution_2004}, who in turn adapted those of \citet{woosley_evolution_1995} for CCSNe and \citet{iwamoto_nucleosynthesis_1999} for SNe Ia to provide a better fit between GCE models and local abundance data. Notably, the yields for O and Fe were left unchanged from the original studies. However, because \citet{woosley_evolution_1995} report gross yields without detailed initial abundances for their CCSN progenitors, and because \citet{francois_evolution_2004} do not provide population-averaged yields, it is difficult to make a comparison with our yield sets. Ultimately, \citet{francois_evolution_2004} report that their GCE models are insensitive to changes in the CCSN yield of O by a factor of 2, so we consider it reasonable to explore the full range of yields given in Table \ref{tab:yields}.

\begin{figure}
    \centering
    \includegraphics{figures/yield_outflow.pdf}
    \caption{The abundance evolution of three one-zone models with different yield sets and outflow mass-loading factors. Table \ref{tab:yields} presents the population-averaged yields for each model. The gray points with error bars indicate the mode of the APOGEE abundance distributions from the Solar neighborhood ($7\leq R_{\rm gal}<9\,{\rm kpc}$, $0\leq|z|<0.5\,{\rm kpc}$) in 1 Gyr-wide bins of the \citet{leung_variational_2023} age estimates.}
    \label{fig:yield-outflow}
\end{figure}

Figure \ref{fig:yield-outflow} illustrates the effect of the yield scaling on the abundance evolution in one-zone models. We vary the outflow mass-loading factor $\eta$ for each model to achieve a consistent endpoint to the abundance evolution (see Section \ref{sec:outflows} for further discussion on outflows). All models feature a rapid dilution of the ISM metallicity by $\sim0.5-0.8$ dex, visible in the top two panels, brought on by the infall of pristine gas at $t_{\rm max}$. For the model with $y/Z_\odot=1$, this dilution persists for some time and the metallicity does not return to Solar until the close to present day. The models with higher yields and outflows recover from this dilution more quickly, returning to Solar metallicity by $\sim5\,{\rm Gyr}$ ago. However, the high-yield models experience a decline in [O/Fe] of $\sim0.2$ dex between the second infall and the present day, contrasted with the smaller decline of $\sim0.1$ dex in the low-yield model.
